---
title: "Draft 1"
author: "Mehrdad Ghyabi, Sai Jaswanth Kunku, Sagar Goswami"
date: "12/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initiation

```{r}
rm(list=ls()) # Dumping memory
```

## Importing Required Libraries

```{r}
library(tidyverse)
library(ggfortify)
library(ggplot2)
library(psych)
```

## Importing Database Files

```{r}
df <- read.csv(file = "Life_Expectancy_Data_cleaned.csv")

df <- df[,-c(1,6,7,9,10,11,12,13,18)]

df$Status <- as.factor(df$Status)
```

------------------------------------------------------------------------

## About the Database

-   <div>

    > -   Country- Country
    >
    > -   Year- Year
    >
    > -   Status- Developed or Developing status
    >
    > -   Life Expectancy- Age(years)
    >
    > -   ~~Adult Mortality- Adult Mortality Rates of both sexes(probability of dying between 15&60 years per 1000 population)~~
    >
    > -   ~~Infant Deaths- Number of Infant Deaths per 1000 population~~
    >
    > -   Alcohol- Alcohol, recorded per capita (15+) consumption (in litres of pure alcohol)
    >
    > -   ~~Percent Expenditure- Expenditure on health as a percentage of Gross Domestic Product per capita(%)~~
    >
    > -   ~~Hep B- Hepatitis B (HepB) immunization coverage among 1-year-olds(%)~~
    >
    > -   ~~Measles- number of reported measles cases per 1000 population~~
    >
    > -   ~~BMI- Average Body Mass Index of entire population~~
    >
    > -   ~~U-5 Deaths- Number of under-five deaths per 1000 population~~
    >
    > -   Polio- Polio(Pol3) immunization coverage among 1-year-olds(%)
    >
    > -   Total Expenditure- General government expenditure on health as a percentage of total government expenditure(%)
    >
    > -   Diphtheria- Diphtheria tetanus toxoid and pertussis (DTP3) immunization coverage among 1-year-olds(%)
    >
    > -   HIV/AIDS- Deaths per 1000 live births HIV/AIDS(0-4 years)
    >
    > -   ~~GDP- Gross Domestic Product per capita(in USD)~~
    >
    > -   Population- Population Thinness 10-19- Prevalence of thinness among children and adolescents for Age 10 to 19(%)
    >
    > -   Thinness 5-9- Prevalence of thinness among children for Age 5 to 9(%)
    >
    > -   Income Composition- Human Development Index in terms of income composition of resources(0-1)
    >
    >     -   filter \> 0
    >
    > -   Schooling- Number of years of Schooling
    >
    >     -   filter \> 1

    </div>

```{r}
df
```

```{r fig.height=10, fig.width=10}
pairs.panels(df)
```

```{r}
summary(df)
```

# Data Cleaning

```{r}
df[rowSums(is.na(df))>0,]
```

### df1: Entire Data

Removing records with more than 5 missing variables.

```{r}
df1 <- df[rowSums(is.na(df))<5,]
```

Removing variables without the Population Data

```{r}
df1 <- df[df$Income.composition.of.resources>0,]
df1 <- df[df$Schooling>0,]
df1 <- df1[!is.na(df1$Population.y),]
df1 <- df1[!is.na(df1$Life.expectancy),]
```

Replacing NA values with the mean values for columns with less than 5% missing values.

```{r}
df1$Alcohol[is.na(df1$Alcohol)] <- mean(df$Alcohol, na.rm = TRUE)
df1$Polio[is.na(df1$Polio)] <- mean(df$Polio, na.rm = TRUE)
df1$Diphtheria[is.na(df1$Diphtheria)] <- mean(df$Diphtheria, na.rm = TRUE)
df1$thinness..1.19.years[is.na(df1$thinness..1.19.years)] <- mean(df$thinness..1.19.years, na.rm = TRUE)
df1$thinness.5.9.years[is.na(df1$thinness.5.9.years)] <- mean(df$thinness.5.9.years, na.rm = TRUE)
df1$Schooling[is.na(df1$Schooling)] <- mean(df$Schooling, na.rm = TRUE)
```

Removing Rows with Null Values

```{r}
df1 <- df1[rowSums(is.na(df1))<1,]
```

### df2014: 2014 Data

```{r}
df2014 <- df[df$Year == 2014,]
df2014
```

```{r}
summary(df2014)
```

### Preparing MapData

```{r}
mapdata <- map_data("world")
names(mapdata)[5] <- "Country"
```

```{r}
Countries_map <- unique(mapdata["Country"])
Countries_map["in_map"] = "map"
Countries_map

Countries_df <- unique(df2014["Country"])
Countries_df["in_df"] = "df"
Countries_df

Countries <- merge(x = Countries_map, y = Countries_df,
                   by = "Country", all = TRUE)
Countries
```

```{r}
df2014[df2014$Country == "Antigua and Barbuda", "Country"] <- "Antigua"
df2014[df2014$Country == "Bolivia (Plurinational State of)", "Country"] <- "Bolivia"
df2014[df2014$Country == "Brunei Darussalam", "Country"] <- "Brunei"
df2014[df2014$Country == "Côte d'Ivoire", "Country"] <- "Ivory Coast"
df2014[df2014$Country == "Cabo Verde", "Country"] <- "Cape Verde"
df2014[df2014$Country == "Congo", "Country"] <- "Republic of Congo"
df2014[df2014$Country == "Czechia", "Country"] <- "Czech Republic"
df2014[df2014$Country == "Democratic People's Republic of Korea", "Country"] <- "North Korea"
df2014[df2014$Country == "Republic of Korea", "Country"] <- "South Korea"
df2014[df2014$Country == "Iran (Islamic Republic of)", "Country"] <- "Iran"
df2014[df2014$Country == "Lao People's Democratic Republic", "Country"] <- "Laos"
df2014[df2014$Country == "Micronesia (Federated States of)", "Country"] <- "Micronesia"
df2014[df2014$Country == "Republic of Moldova", "Country"] <- "Moldova"
df2014[df2014$Country == "Russian Federation", "Country"] <- "Russia"
df2014[df2014$Country == "Saint Kitts and Nevis", "Country"] <- "Saint Kitts"
df2014[df2014$Country == "Saint Vincent and the Grenadines", "Country"] <- "Saint Vincent"
df2014[df2014$Country == "Syrian Arab Republic", "Country"] <- "Syria"
df2014[df2014$Country == "The former Yugoslav republic of Macedonia", "Country"] <- "Macedonia"
df2014[df2014$Country == "Trinidad and Tobago", "Country"] <- "Trinidad"
df2014[df2014$Country == "United Kingdom of Great Britain and Northern Ireland", "Country"] <- "UK"
df2014[df2014$Country == "United Republic of Tanzania", "Country"] <- "Tanzania"
df2014[df2014$Country == "United States of America", "Country"] <- "USA"
df2014[df2014$Country == "Venezuela (Bolivarian Republic of)", "Country"] <- "Venezuela"
df2014[df2014$Country == "Viet Nam", "Country"] <- "Vietnam"

#df2014[df2014$Country == "xxxxxxxxxxxxxxxxxx", "Country"] <- "xxxxxxxxxxxxxxxxxx"
```

```{r}
mapdata <- left_join(mapdata, df2014, by = "Country")
```

# Data Exploration

## Univariate Analysis \| Understanding the Raw Data (Map Visualizations)

#### Life Expectancy

```{r fig.height=6, fig.width=10}
MapLifeExpectancy <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = Life.expectancy),
                     color = "black") +
        labs(title = "Life Expectancy [Years]",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_distiller(palette = "Spectral", direction = 1) +
        theme_void()
MapLifeExpectancy
```

```{r}
ggplot(df) +
    aes(x = Year, y = Life.expectancy, colour = Status) +
    geom_point(shape = "bullet", size = 0.1, alpha = 0.25) +
    geom_smooth(span = 0.75) +
    scale_color_brewer(palette = "Set2", direction = 1) +
    labs(x = "Year",
         y = "Life Expectancy [in Years]",
         title = "Change in Life Expectancy over the Years",
         caption = "Data Source: World Health Organization") +
    theme_bw()
```

#### Alcohol Consumption

```{r fig.height=6, fig.width=10}
MapAlcohol <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = Alcohol),
                     color = "black") +
        labs(title = "Alcohol Consumption per capita [Litres of Pure Alcohol]",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_distiller(palette = "Spectral", direction = -1) +
        theme_void()
MapAlcohol
```

#### Polio Vaccinations

```{r fig.height=6, fig.width=10}
MapPolio <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = Polio),
                     color = "black") +
        labs(title = "% Polio Vaccination (Pol3) among Children",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_distiller(palette = "Spectral", direction = 1) +
        theme_void()
MapPolio
```

#### Total Expenditure

```{r fig.height=6, fig.width=10}
MapTotal.expenditure <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = Total.expenditure),
                     color = "black") +
        labs(title = "General Government Expenditure on Health \n[% of Total Govt. Expenditure]",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_distiller(palette = "Spectral", direction = 1) +
        theme_void()
MapTotal.expenditure
```

#### Diphtheria

```{r fig.height=6, fig.width=10}
MapDiphtheria <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = Diphtheria),
                     color = "black") +
        labs(title = "Diphtheria/Tetanus/Pertussis (DTP3) Immunization Coverage among Children",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_distiller(palette = "Spectral", direction = 1) +
        theme_void()
MapDiphtheria
```

#### HIV AIDS

```{r fig.height=6, fig.width=10}
MapHivAids <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = HIV.AIDS),
                     color = "black") +
        labs(title = "Deaths per 1000 live-births due to HIV/AIDS",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_distiller(palette = "Spectral", direction = -1) +
        theme_void()
MapHivAids
```

#### Thinness in Children

```{r fig.height=6, fig.width=10}
MapThinness <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = thinness..1.19.years),
                     color = "black") +
        labs(title = "% Prevalence of Thinness among Children",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_distiller(palette = "Spectral", direction = -1) +
        theme_void()
MapThinness
```

#### Human Development Index

```{r fig.height=6, fig.width=10}
MapHDI <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = Income.composition.of.resources),
                     color = "black") +
        labs(title = "Human Development Index as \nIncome Composition of Resources [0-1]",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_distiller(palette = "Spectral", direction = 1) +
        theme_void()
MapHDI
```

#### Schooling

```{r fig.height=6, fig.width=10}
MapSchooling <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = Schooling),
                     color = "black") +
        labs(title = "Schooling [No of Years]",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_distiller(palette = "Spectral", direction = 1) +
        theme_void()
MapSchooling
```

## Bi-variate Analysis

```{r fig.height=9, fig.width=16}
pairs.panels(df)
```

```{r}
matrix(colnames(df))
```

### Relationship between Government Health Expenditure and Life Expectancy

```{r fig.height=6, fig.width=10}
df %>%
    
    ggplot() +
    aes(x = Total.expenditure, y = Life.expectancy, colour = Status) +
    
    geom_point(shape = "circle", size = 1.5, alpha = 0.2) +
    geom_smooth(method = "lm") +
    
    scale_color_manual(values = c(Developed = "#1B9E77",
                                  Developing = "#666666")) +
    
    labs(x = "Government Health Expenditure [% of Total Expenditure]",
         y = "Life Expectancy [in Years]",
         title = "Relationship between Government Health Expenditure and Life Expectancy",
         caption = "Data Source: World Health Organization") +
    
    theme_bw()
```

### Relationship between Alcohol Consumption and Life Expectancy

```{r fig.height=6, fig.width=10}
df %>%
    
    ggplot() +
    aes(x = Alcohol, y = Life.expectancy, colour = Status) +
    
    geom_point(shape = "circle", size = 1.5, alpha = 0.2) +
    geom_smooth() +
    
    scale_color_manual(values = c(Developed = "#1B9E77",
                                  Developing = "#666666")) +
    
    labs(x = "Alcohol Consumption per Capita [Lts. of Pure Alcohol]",
         y = "Life Expectancy [in Years]",
         title = "Relationship between Alcohol Consumption and Life Expectancy",
         caption = "Data Source: World Health Organization") +
    
    theme_bw()
```

### Relationship between Thinness in Children and Life Expectancy

```{r fig.height=6, fig.width=10}
df %>%
    
    ggplot() +
    aes(x = thinness..1.19.years, y = Life.expectancy, colour = Status) +
    
    geom_point(shape = "circle", size = 1.5, alpha = 0.2) +
    geom_smooth(method = "lm") +
    
    scale_color_manual(values = c(Developed = "#1B9E77",
                                  Developing = "#666666")) +
    
    labs(x = "Thinness in Children [% of Population]",
         y = "Life Expectancy [in Years]",
         title = "Relationship between Thinness in Children and Life Expectancy",
         caption = "Data Source: World Health Organization") +
    
    theme_bw()
```

### Relationship between Schooling and Life Expectancy

```{r fig.height=6, fig.width=10}
df %>%
    filter(Schooling >= 0.1 | is.na(Schooling)) %>%
    
    ggplot() +
    aes(x = Schooling, y = Life.expectancy, colour = Status) +
    
    geom_point(shape = "circle", size = 1.5, alpha = 0.2) +
    geom_smooth(method = "lm") +
    
    scale_color_manual(values = c(Developed = "#1B9E77",
                                  Developing = "#666666")) +
    
    labs(x = "Schooling [in Years]",
         y = "Life Expectancy [in Years]",
         title = "Relationship between Schooling and Life Expectancy",
         caption = "Data Source: World Health Organization") +
    
    theme_bw()
```

### Relationship between Schooling and Human Development Index

```{r fig.height=6, fig.width=10}
df %>%
    filter(Schooling >= 0.1 | is.na(Schooling)) %>%
    filter(Income.composition.of.resources >= 0.1 | is.na(Income.composition.of.resources)) %>%
    
    ggplot() +
    aes(x = Schooling, y = Income.composition.of.resources, colour = Status) +
    
    geom_point(shape = "circle", size = 1.5, alpha = 0.2) +
    geom_smooth(method = "lm") +
    
    scale_color_manual(values = c(Developed = "#1B9E77",
                                  Developing = "#666666")) +
    
    labs(x = "Schooling [in Years]",
         y = "Human Development Index [0-1]",
         title = "Relationship between Schooling and Human Development Index",
         caption = "Data Source: World Health Organization") +
    
    theme_bw()
```

### Relationship between Vaccinations and Life Expectancy

```{r fig.height=6, fig.width=10}
df %>%
    filter(Polio >= 10 | is.na(Polio)) %>%
    
    ggplot() +
    aes(x = Polio, y = Life.expectancy, colour = Status) +
    
    geom_point(shape = "circle", size = 1.5, alpha = 0.2) +
    geom_smooth(method = "lm") +
    
    scale_color_manual(values = c(Developed = "#1B9E77",
                                  Developing = "#666666")) +
    
    labs(x = "Polio Vaccination [% of Children]",
         y = "Life Expectancy [in Years]",
         title = "Relationship between Polio Vaccination and Life Expectancy",
         caption = "Data Source: World Health Organization") +
    
    theme_bw()
```

```{r fig.height=6, fig.width=10}
df %>%
    filter(Diphtheria >= 10 | is.na(Diphtheria)) %>%
    
    ggplot() +
    aes(x = Diphtheria, y = Life.expectancy, colour = Status) +
    
    geom_point(shape = "circle", size = 1.5, alpha = 0.2) +
    geom_smooth(method = "lm") +
    
    scale_color_manual(values = c(Developed = "#1B9E77",
                                  Developing = "#666666")) +
    
    labs(x = "Diphtheria Vaccination [% of Children]",
         y = "Life Expectancy [in Years]",
         title = "Relationship between Diphtheria Vaccination and Life Expectancy",
         caption = "Data Source: World Health Organization") +
    
    theme_bw()
```

## Data Clusters

#### Creating Clusters

```{r}
df2014_trim <- df2014[rowSums(is.na(df2014)) == 0,]
```

```{r}
k = 4
km <- kmeans(df2014_trim[,c(4,6,8,9,13)], centers=k)

df2014_trim$cluster <- as.factor(km$cluster)

autoplot(km, df2014_trim[,4:14], frame = TRUE)
```

#### Preparing MapData

```{r}
mapdata <- map_data("world")
names(mapdata)[5] <- "Country"
```

```{r}
mapdata <- left_join(mapdata, df2014_trim, by = "Country")
```

#### Mapping Clusters

```{r}
MapClusters <- ggplot(data = mapdata,
               aes(x = long, y = lat, group = group)) +
        geom_polygon(aes(fill = cluster),
                     color = "black") +
        labs(title = "Clustering Health Data",
            caption=c('Data Source: World Health Organization')) +
        scale_fill_brewer(name = "Clusters", palette = "Pastel2") +
        theme_void()
MapClusters
```

```{r}
print("Done...")
```
